---
title: "Cholla climate-demography relationships"
author: "Tom Miller"
date: "July 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message=F,warning=F)
library(tidyverse)
library(lme4)
library(mcmcplots)
library(R2jags)
PCclim <- read.csv("Analysis with ClimateWNA\\ClimateWNA PC Values.csv")
volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit<-function(x){exp(x)/(1+exp(x))}
# there are the binning parameters for visualization
PC_bins <- 4
size_bin_num <- 3
```

# Purpose: exploration of historical and future SEV climate and fitting climate-dependent demography functions for IPM analysis

## How much over lap do we have across the 3 climate groups (past, present, future)?
```{r PC hist}
PC_gather <- PCclim %>% 
  select(-PC4,-PC5,-PC6,-PC7,-PC8) %>% 
  gather(PC1,PC2,PC3,key="PC",value="value") %>% 
  mutate(time = ifelse(Year_t<2004,"past",ifelse(Year_t>2016,"future","present")))

ggplot(PC_gather)+
  geom_histogram(aes(x=value,fill=time))+
    facet_grid(PC~.)
```
The plot above is useful to see and a little bit concerning. In order to project past and future responses to climate, we will need to extrapolate our statistical models well beyond the limits of observed values, especially for PC1. Let's see what this looks like in real time. 
```{r PC timeseries}
ggplot(PC_gather)+
  geom_line(aes(x=Year_t,y=value,color=time))+
    facet_grid(PC~.)
```

I mean, I guess this is what climate change looks like. I think this argues for testing polynomial responses - but even that is assuming a particular type of extrapolation that could be equally wrong. Let's see how the parameter estimates shape up. 

## Demography
```{r read cholla dat}
cholla <- read.csv("cholla_demography_20042018.csv")
## merge demography and climate,
## subset to demography years only,
## drop seed addition plots (don't want them in plot RFX)
## change plots and years to 1,...,N integers

cholla.clim <- full_join(cholla,PCclim,by="Year_t") %>% 
  filter(Year_t >= min(cholla$Year_t,na.rm=T),
         Year_t <= max(cholla$Year_t,na.rm=T)) %>% 
  filter(str_sub(Plot,1,1)!="H") %>% 
  mutate(year_int = Year_t - (min(Year_t)-1),
       plot_int = as.integer(Plot),
       vol_t = log(volume(h = Height_t, w = Width_t, p = Perp_t)),
       vol_t1 = log(volume(h = Height_t1, w = Width_t1, p = Perp_t1)),
       standvol_t = (vol_t - mean(vol_t,na.rm=T))/sd(vol_t,na.rm=T),
       standvol_t1 = (vol_t1 - mean(vol_t1,na.rm=T))/sd(vol_t1,na.rm=T))

```

### 1. Flowering
```{r miscellany data prep for flowering}
## cut NAs and bundle data
flow_dat <- cholla.clim %>% 
  filter(!is.na(standvol_t),
         !is.na(Goodbuds_t))

PC_range <- PC_gather%>% 
  summarise(PC_min = min(value),
            PC_max = max(value))
x_PC = seq(PC_range$PC_min,PC_range$PC_max,0.1)

# not sure if I want to bin by equal intervals or equal sample sizes. Might explore both.
flow_dat_cuts <- flow_dat %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num))
size_cuts <- flow_dat_cuts %>% 
  group_by(size_bin) %>% 
  summarise(bin_mean = mean(standvol_t))
hist(flow_dat$standvol_t)
abline(v = size_cuts$bin_mean)
```

I am starting the demographic analysis with a vital rate that I know to be variable and likely climate-dependent. This should be relatively easy to fit and we'll do the hard stuff later. 

```{r flow fit}
flow.dat<-list(N.plots = max(flow_dat$plot_int),
               N.years = max(flow_dat$year_int),
               N.obs = nrow(flow_dat),
               plot = flow_dat$plot_int,
               year = flow_dat$year_int,
               PC1 = flow_dat$PC1,
               PC2 = flow_dat$PC2,
               PC3 = flow_dat$PC3,
               size = flow_dat$standvol_t,
               y = flow_dat$Goodbuds_t > 0,
               x_PC = x_PC,
               PC_levels = length(x_PC),
               x_size = size_cuts$bin_mean,
               size_levels = length(size_cuts$bin_mean))

## Inits function
inits<-function(){list(mu.alpha=rnorm(1),
                       betasize=rnorm(1),
                       betaPC1=rnorm(1),
                       betaPC1q=rnorm(1),
                       betaPC1_size=rnorm(1),
                       betaPC1q_size=rnorm(1),
                       betaPC2=rnorm(1),
                       betaPC2q=rnorm(1),
                       betaPC2_size=rnorm(1),
                       betaPC2q_size=rnorm(1),
                       betaPC3=rnorm(1),
                       betaPC3q=rnorm(1),
                       betaPC3_size=rnorm(1),
                       betaPC3q_size=rnorm(1),
                       sigma.plot=rlnorm(1),
                       sigma.year=rlnorm(1))}

## Params to estimate
parameters<-c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.plot","sigma.year",
              "fit","fit.new")

## MCMC settings
ni<-10000
nb<-1000
nt<-5
nc<-3

```

```{r run JAGS model}
## run JAGS
cholla.flow.out<-jags(data=flow.dat,inits=inits,parameters.to.save=parameters,model.file="Bayes models\\HB_cholla_flowering.txt",
          n.thin=nt,n.chains=nc,n.burnin=nb,n.iter=ni,DIC=T,working.directory=getwd())
```

```{r flower model diagnostics}
mcmcplot(cholla.flow.out, parms = c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.plot","sigma.year"))

plot(cholla.flow.out$BUGSoutput$sims.list$fit,
     cholla.flow.out$BUGSoutput$sims.list$fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
```

Fits and convergence look good. Let's visualize the results.
```{r visualize results}

flow_PC1_cuts <- flow_dat_cuts%>% 
  mutate(PC1_bin = cut_interval(PC1,PC_bins),
            size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,PC1_bin) %>% 
  summarise(mean_PC1 = mean(PC1),
            mean_flow = mean(Goodbuds_t > 0),
            bin_n = n())
flow_PC2_cuts <- flow_dat_cuts%>% 
  mutate(PC2_bin = cut_interval(PC2,PC_bins),
            size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,PC2_bin) %>% 
  summarise(mean_PC2 = mean(PC2),
            mean_flow = mean(Goodbuds_t > 0),
            bin_n = n())
flow_PC3_cuts <- flow_dat_cuts%>% 
  mutate(PC3_bin = cut_interval(PC3,PC_bins),
            size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,PC3_bin) %>% 
  summarise(mean_PC3 = mean(PC3),
            mean_flow = mean(Goodbuds_t > 0),
            bin_n = n())

mean_params <- cholla.flow.out$BUGSoutput$mean

par(mfrow=c(1,3))
with(flow_PC1_cuts,{plot(mean_PC1,mean_flow,col=size_bins,pch=16,cex= 3*(bin_n / max(bin_n)),ylim=c(0,1))})
for(i in 1:length(size_cuts$bin_mean)){
lines(x_PC,
      invlogit(mean_params$mu.alpha + mean_params$betaPC1 * x_PC + mean_params$betaPC1q * x_PC^2 +
      mean_params$betasize * size_cuts$bin_mean[i] + mean_params$betaPC1_size * x_PC * size_cuts$bin_mean[i] +
        mean_params$betaPC1q_size * (x_PC^2) * size_cuts$bin_mean[i]),
      lwd=2,col=i)
}

with(flow_PC2_cuts,{plot(mean_PC2,mean_flow,col=size_bins,pch=16,cex= 3*(bin_n / max(bin_n)),ylim=c(0,1))})
for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC,
      invlogit(mean_params$mu.alpha + mean_params$betaPC2 * x_PC + mean_params$betaPC2q * x_PC^2 +
      mean_params$betasize * size_cuts$bin_mean[i] + mean_params$betaPC2_size * x_PC * size_cuts$bin_mean[i] +
        mean_params$betaPC2q_size * (x_PC^2) * size_cuts$bin_mean[i]),
      lwd=2,col=i)
}

with(flow_PC3_cuts,{plot(mean_PC3,mean_flow,col=size_bins,pch=16,cex= 3*(bin_n / max(bin_n)),ylim=c(0,1))})
for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC,
      invlogit(mean_params$mu.alpha + mean_params$betaPC3 * x_PC + mean_params$betaPC3q * x_PC^2 +
      mean_params$betasize * size_cuts$bin_mean[i] + mean_params$betaPC3_size * x_PC * size_cuts$bin_mean[i] +
        mean_params$betaPC3q_size * (x_PC^2) * size_cuts$bin_mean[i]),
      lwd=2,col=i)
}

```

This looks good! Moving on to survival.

### 2. Survival
```{r miscellany data prep for survival}
## cut NAs and bundle data
surv_dat <- cholla.clim %>% 
  filter(!is.na(standvol_t),
         !is.na(Survival_t1))

# not sure if I want to bin by equal intervals or equal sample sizes. Might explore both.
surv_dat_cuts <- surv_dat %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num))
surv_size_cuts <- surv_dat_cuts %>% 
  group_by(size_bin) %>% 
  summarise(bin_mean = mean(standvol_t))
hist(surv_dat$standvol_t)
abline(v = surv_size_cuts$bin_mean)
```

```{r surv fit}
surv.dat<-list(N.plots = max(surv_dat$plot_int),
               N.years = max(surv_dat$year_int),
               N.obs = nrow(surv_dat),
               plot = surv_dat$plot_int,
               year = surv_dat$year_int,
               PC1 = surv_dat$PC1,
               PC2 = surv_dat$PC2,
               PC3 = surv_dat$PC3,
               size = surv_dat$standvol_t,
               y = surv_dat$Survival_t1,
               x_PC = x_PC,
               PC_levels = length(x_PC),
               x_size = surv_size_cuts$bin_mean,
               size_levels = length(surv_size_cuts$bin_mean))

## Inits function
inits<-function(){list(mu.alpha=rnorm(1),
                       betasize=rnorm(1),
                       betaPC1=rnorm(1),
                       betaPC1q=rnorm(1),
                       betaPC1_size=rnorm(1),
                       betaPC1q_size=rnorm(1),
                       betaPC2=rnorm(1),
                       betaPC2q=rnorm(1),
                       betaPC2_size=rnorm(1),
                       betaPC2q_size=rnorm(1),
                       betaPC3=rnorm(1),
                       betaPC3q=rnorm(1),
                       betaPC3_size=rnorm(1),
                       betaPC3q_size=rnorm(1),
                       sigma.plot=rlnorm(1),
                       sigma.year=rlnorm(1))}

## Params to estimate
parameters<-c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.plot","sigma.year",
              "fit","fit.new")

## MCMC settings
ni<-10000
nb<-1000
nt<-5
nc<-3

```

```{r run JAGS model}
## run JAGS
cholla.surv.out<-jags(data=surv.dat,inits=inits,parameters.to.save=parameters,model.file="Bayes models\\HB_cholla_survival.txt",
          n.thin=nt,n.chains=nc,n.burnin=nb,n.iter=ni,DIC=T,working.directory=getwd())
```

```{r survival model diagnostics}
mcmcplot(cholla.surv.out, parms = c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.plot","sigma.year"))

plot(cholla.surv.out$BUGSoutput$sims.list$fit,
     cholla.surv.out$BUGSoutput$sims.list$fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
```


```{r visualize survival results}

surv_PC1_cuts <- surv_dat_cuts%>% 
  mutate(PC1_bin = cut_interval(PC1,PC_bins),
            size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,PC1_bin) %>% 
  summarise(mean_PC1 = mean(PC1),
            mean_surv = mean(Survival_t1),
            bin_n = n())
surv_PC2_cuts <- surv_dat_cuts%>% 
  mutate(PC2_bin = cut_interval(PC2,PC_bins),
            size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,PC2_bin) %>% 
  summarise(mean_PC2 = mean(PC2),
            mean_surv = mean(Survival_t1),
            bin_n = n())
surv_PC3_cuts <- surv_dat_cuts%>% 
  mutate(PC3_bin = cut_interval(PC3,PC_bins),
            size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,PC3_bin) %>% 
  summarise(mean_PC3 = mean(PC3),
            mean_surv = mean(Survival_t1),
            bin_n = n())

mean_surv_params <- cholla.surv.out$BUGSoutput$mean

par(mfrow=c(1,3))
with(surv_PC1_cuts,{plot(mean_PC1,mean_surv,col=size_bins,pch=16,ylim=c(0,1),cex= 3*(bin_n / max(bin_n)))})
for(i in 1:length(surv_size_cuts$bin_mean)){
lines(x_PC,
      invlogit(mean_surv_params$mu.alpha + mean_surv_params$betaPC1 * x_PC + mean_surv_params$betaPC1q * x_PC^2 +
      mean_surv_params$betasize * surv_size_cuts$bin_mean[i] + mean_surv_params$betaPC1_size * x_PC * surv_size_cuts$bin_mean[i] +
        mean_surv_params$betaPC1q_size * (x_PC^2) * surv_size_cuts$bin_mean[i]),
      lwd=2,col=i)
}

with(surv_PC2_cuts,{plot(mean_PC2,mean_surv,col=size_bins,pch=16,cex= 3*(bin_n / max(bin_n)),ylim=c(0,1))})
for(i in 1:length(surv_size_cuts$bin_mean)){
  lines(x_PC,
      invlogit(mean_surv_params$mu.alpha + mean_surv_params$betaPC2 * x_PC + mean_surv_params$betaPC2q * x_PC^2 +
      mean_surv_params$betasize * surv_size_cuts$bin_mean[i] + mean_surv_params$betaPC2_size * x_PC * surv_size_cuts$bin_mean[i] +
        mean_surv_params$betaPC2q_size * (x_PC^2) * surv_size_cuts$bin_mean[i]),
      lwd=2,col=i)
}

with(surv_PC3_cuts,{plot(mean_PC3,mean_surv,col=size_bins,pch=16,cex= 3*(bin_n / max(bin_n)),ylim=c(0,1))})
for(i in 1:length(surv_size_cuts$bin_mean)){
  lines(x_PC,
      invlogit(mean_surv_params$mu.alpha + mean_surv_params$betaPC3 * x_PC + mean_surv_params$betaPC3q * x_PC^2 +
      mean_surv_params$betasize * surv_size_cuts$bin_mean[i] + mean_surv_params$betaPC3_size * x_PC * surv_size_cuts$bin_mean[i] +
        mean_surv_params$betaPC3q_size * (x_PC^2) * surv_size_cuts$bin_mean[i]),
      lwd=2,col=i)
}

```

Hmmm. Not sure how I feel about these fits. 

##3. Growth
```{r grow_dat}
## cut NAs and bundle data
grow_dat <- cholla.clim %>% 
  filter(!is.na(standvol_t),
         !is.na(standvol_t1)) %>% 
  mutate(growth_t1 = standvol_t1 - standvol_t)

grow.dat<-list(N.plots = max(grow_dat$plot_int),
               N.years = max(grow_dat$year_int),
               N.obs = nrow(grow_dat),
               plot = grow_dat$plot_int,
               year = grow_dat$year_int,
               PC1 = grow_dat$PC1,
               PC2 = grow_dat$PC2,
               PC3 = grow_dat$PC3,
               size = grow_dat$standvol_t,
               y = grow_dat$growth_t1)

## Inits function
inits<-function(){list(mu.alpha=rnorm(1),
                       betasize=rnorm(1),
                       betaPC1=rnorm(1),
                       betaPC1q=rnorm(1),
                       betaPC1_size=rnorm(1),
                       betaPC1q_size=rnorm(1),
                       betaPC2=rnorm(1),
                       betaPC2q=rnorm(1),
                       betaPC2_size=rnorm(1),
                       betaPC2q_size=rnorm(1),
                       betaPC3=rnorm(1),
                       betaPC3q=rnorm(1),
                       betaPC3_size=rnorm(1),
                       betaPC3q_size=rnorm(1),
                       sigma.eps=rlnorm(1),
                       sigma.plot=rlnorm(1),
                       sigma.year=rlnorm(1))}

## Params to estimate
parameters<-c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.eps","sigma.plot","sigma.year",
              "fit","fit.new")

## MCMC settings
ni<-10000
nb<-1000
nt<-5
nc<-3

```

```{r run growth model}
## run JAGS
cholla.grow.out<-jags(data=grow.dat,inits=inits,parameters.to.save=parameters,model.file="Bayes models\\HB_cholla_growth.txt",
          n.thin=nt,n.chains=nc,n.burnin=nb,n.iter=ni,DIC=T,working.directory=getwd())
```

```{r growth model diagnostics}
mcmcplot(cholla.grow.out, parms = c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.plot","sigma.year","sigma.eps"))

plot(cholla.grow.out$BUGSoutput$sims.list$fit,
     cholla.grow.out$BUGSoutput$sims.list$fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
```
```{r visualize growth results}
grow_dat_cuts <- grow_dat %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num),
         size_bins = as.integer(size_bin)) 
grow_size_cuts <- grow_dat_cuts %>% 
  group_by(size_bin) %>% 
  summarise(bin_mean = mean(standvol_t))
  
mean_grow_params <- cholla.grow.out$BUGSoutput$mean

par(mfrow=c(1,3))
with(grow_dat_cuts,{plot(PC1,growth_t1,col=alpha(size_bins,0.25),pch=1,cex= 1)})
for(i in 1:length(grow_size_cuts$bin_mean)){
lines(x_PC,
      mean_grow_params$mu.alpha + mean_grow_params$betaPC1 * x_PC + mean_grow_params$betaPC1q * x_PC^2 +
      mean_grow_params$betasize * grow_size_cuts$bin_mean[i] + mean_grow_params$betaPC1_size * x_PC * grow_size_cuts$bin_mean[i] +
        mean_grow_params$betaPC1q_size * (x_PC^2) * grow_size_cuts$bin_mean[i],
      lwd=2,col=i)
}

with(grow_dat_cuts,{plot(PC2,growth_t1,col=alpha(size_bins,0.25),pch=1,cex= 1)})
for(i in 1:length(grow_size_cuts$bin_mean)){
lines(x_PC,
      mean_grow_params$mu.alpha + mean_grow_params$betaPC2 * x_PC + mean_grow_params$betaPC2q * x_PC^2 +
      mean_grow_params$betasize * grow_size_cuts$bin_mean[i] + mean_grow_params$betaPC2_size * x_PC * grow_size_cuts$bin_mean[i] +
        mean_grow_params$betaPC2q_size * (x_PC^2) * grow_size_cuts$bin_mean[i],
      lwd=2,col=i)
}

with(grow_dat_cuts,{plot(PC3,growth_t1,col=alpha(size_bins,0.25),pch=1,cex= 1)})
for(i in 1:length(grow_size_cuts$bin_mean)){
lines(x_PC,
      mean_grow_params$mu.alpha + mean_grow_params$betaPC3 * x_PC + mean_grow_params$betaPC3q * x_PC^2 +
      mean_grow_params$betasize * grow_size_cuts$bin_mean[i] + mean_grow_params$betaPC3_size * x_PC * grow_size_cuts$bin_mean[i] +
        mean_grow_params$betaPC3q_size * (x_PC^2) * grow_size_cuts$bin_mean[i],
      lwd=2,col=i)
}


```


##4. Fertility
```{r grow_dat}
## cut NAs and bundle data
fert_dat <- cholla.clim %>% 
  filter(Goodbuds_t>0,
         !is.na(standvol_t),
         !is.na(Goodbuds_t))

fert.dat<-list(N.plots = max(fert_dat$plot_int),
               N.years = max(fert_dat$year_int),
               N.obs = nrow(fert_dat),
               plot = fert_dat$plot_int,
               year = fert_dat$year_int,
               PC1 = fert_dat$PC1,
               PC2 = fert_dat$PC2,
               PC3 = fert_dat$PC3,
               size = fert_dat$standvol_t,
               y = fert_dat$Goodbuds_t)

## Inits function
inits<-function(){list(mu.alpha=rnorm(1),
                       betasize=rnorm(1),
                       betaPC1=rnorm(1),
                       betaPC1q=rnorm(1),
                       betaPC1_size=rnorm(1),
                       betaPC1q_size=rnorm(1),
                       betaPC2=rnorm(1),
                       betaPC2q=rnorm(1),
                       betaPC2_size=rnorm(1),
                       betaPC2q_size=rnorm(1),
                       betaPC3=rnorm(1),
                       betaPC3q=rnorm(1),
                       betaPC3_size=rnorm(1),
                       betaPC3q_size=rnorm(1),
                       sigma.plot=rlnorm(1),
                       sigma.year=rlnorm(1))}

## Params to estimate
parameters<-c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.plot","sigma.year",
              "fit","fit.new")

## MCMC settings
ni<-10000
nb<-1000
nt<-5
nc<-3

```

```{r run fert model}
## run JAGS
cholla.fert.out<-jags(data=fert.dat,inits=inits,parameters.to.save=parameters,
                      model.file="Bayes models\\HB_cholla_fertility.txt",
          n.thin=nt,n.chains=nc,n.burnin=nb,n.iter=ni,DIC=T,working.directory=getwd())
```

```{r fert model diagnostics}
mcmcplot(cholla.fert.out, parms = c("mu.alpha","betasize",
              "betaPC1","betaPC1q","betaPC1_size","betaPC1q_size",
              "betaPC2","betaPC2q","betaPC2_size","betaPC2q_size",
              "betaPC3","betaPC3q","betaPC3_size","betaPC3q_size",
              "sigma.plot","sigma.year"))

plot(cholla.fert.out$BUGSoutput$sims.list$fit,
     cholla.fert.out$BUGSoutput$sims.list$fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
```













