---
title: "Cholla climate-demography relationships"
author: "Tom Miller"
date: "September 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message=F,warning=F)
library(tidyverse)
library(lme4)
library(mcmcplots)
library(R2jags)
PCclim <- read.csv("ClimateWNA_PCvalues_out.csv")
volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit<-function(x){exp(x)/(1+exp(x))}
```

# Purpose: exploration of historical and future SEV climate and fitting climate-dependent demography functions for IPM analysis

## How much over lap do we have between historical data and observation period?
```{r PC hist}
PC_gather <- PCclim %>% 
  select(-PC4,-PC5,-PC6,-PC7,-PC8) %>% 
  gather(PC1,PC2,PC3,key="PC",value="value") %>% 
  mutate(time = ifelse(Year_t<2004,"historical",ifelse(Year_t>2016,"future","observation period")))

PC_range <- PC_gather %>% 
  group_by(PC) %>% 
  summarise(PC_min = min(value),
            PC_max = max(value))
x_PC1 = seq(PC_range$PC_min[PC_range$PC=="PC1"],PC_range$PC_max[PC_range$PC=="PC1"],0.1)
x_PC2 = seq(PC_range$PC_min[PC_range$PC=="PC2"],PC_range$PC_max[PC_range$PC=="PC2"],0.1)
x_PC3 = seq(PC_range$PC_min[PC_range$PC=="PC3"],PC_range$PC_max[PC_range$PC=="PC3"],0.1)

ggplot(PC_gather)+
  geom_histogram(aes(x=value,fill=time))+
    facet_grid(PC~.)
```
The plot above is useful to see and a little bit concerning. We will need to extrapolate demographic performance to high values of PC1 (not observed in demographic study), which I think correspond to cool, big monsoon years. Let's see what this looks like in real time. 
```{r PC timeseries}
ggplot(PC_gather)+
  geom_line(aes(x=Year_t,y=value,color=time))+
    facet_grid(PC~.)
```

I mean, I guess this is what climate change looks like. I think this argues for testing polynomial responses - but even that is assuming a particular type of extrapolation that could be equally wrong. Let's see how the parameter estimates shape up. 

## Demographic data
Read in cholla data and merge in with the climate PCs. Also perform some operations to get data in the format we need.
```{r read cholla dat}
cholla <- read.csv("cholla_demography_20042018.csv")
## merge demography and climate,
## subset to demography years only,
## drop seed addition plots (don't want them in plot RFX)
## change plots and years to 1,...,N integers
cholla.clim <- full_join(cholla,PCclim,by="Year_t") %>% 
  filter(Year_t >= min(cholla$Year_t,na.rm=T),
         Year_t <= max(cholla$Year_t,na.rm=T)) %>% 
  filter(str_sub(Plot,1,1)!="H") %>% 
  mutate(year_int = Year_t - (min(Year_t)-1),
       plot_int = as.integer(Plot),
       vol_t = log(volume(h = Height_t, w = Width_t, p = Perp_t)),
       vol_t1 = log(volume(h = Height_t1, w = Width_t1, p = Perp_t1)),
       standvol_t = (vol_t - mean(vol_t,na.rm=T))/sd(vol_t,na.rm=T),
       standvol_t1 = (vol_t1 - mean(vol_t1,na.rm=T))/sd(vol_t1,na.rm=T))
```

### Bayesian model fitting. 
All vital rates will share the same linear predictor, including size, three climate PCs, quadratic effects of PCs, and all interactions with size (but no PC*PC interactions):

$$\mu = \beta_0+\beta_1*size+
\beta_2*PC1+\beta_3*PC1^{2}+\beta_4*size*PC1+\beta_5*size*PC1^{2}+\rho+\tau$$

Plus the same terms for PC2 and PC3 (not shown for space), so a total of 14 fixed-effect coefficient. Also there are normally distributed random deviates for plot ($\rho$) and year ($\tau$). The JAGS models are written in separate .txt files. 

```{r data prep}
## cut NAs and bundle data
flow_dat <- cholla.clim %>% 
  filter(!is.na(standvol_t),
         !is.na(Goodbuds_t),
         !is.na(PC1))
surv_dat <- cholla.clim %>% 
  filter(!is.na(standvol_t),
         !is.na(Survival_t1),
         !is.na(PC1))
grow_dat <- cholla.clim %>% 
  filter(!is.na(standvol_t),
         !is.na(standvol_t1),
         !is.na(PC1)) %>% 
  mutate(growth_t1 = standvol_t1 - standvol_t)
fert_dat <- cholla.clim %>% 
  filter(Goodbuds_t>0,
         !is.na(standvol_t),
         !is.na(Goodbuds_t),
         !is.na(PC1))

## N plots and years should be the same across vital rates, just check
max(flow_dat$year_int);max(fert_dat$year_int);max(grow_dat$year_int);max(surv_dat$year_int)
max(flow_dat$plot_int);max(fert_dat$plot_int);max(grow_dat$plot_int);max(surv_dat$plot_int)

cholla.dat<-list(N.plots = max(flow_dat$plot_int),
               N.years = max(flow_dat$year_int),
               
               flow.N.obs = nrow(flow_dat),
               flow.plot = flow_dat$plot_int,
               flow.year = flow_dat$year_int,
               flow.PC1 = flow_dat$PC1,
               flow.PC2 = flow_dat$PC2,
               flow.PC3 = flow_dat$PC3,
               flow.size = flow_dat$standvol_t,
               flow.y = flow_dat$Goodbuds_t > 0,
               
               surv.N.obs = nrow(surv_dat),
               surv.plot = surv_dat$plot_int,
               surv.year = surv_dat$year_int,
               surv.PC1 = surv_dat$PC1,
               surv.PC2 = surv_dat$PC2,
               surv.PC3 = surv_dat$PC3,
               surv.size = surv_dat$standvol_t,
               surv.y = surv_dat$Survival_t1,
               
               grow.N.obs = nrow(grow_dat),
               grow.plot = grow_dat$plot_int,
               grow.year = grow_dat$year_int,
               grow.PC1 = grow_dat$PC1,
               grow.PC2 = grow_dat$PC2,
               grow.PC3 = grow_dat$PC3,
               grow.size = grow_dat$standvol_t,
               grow.y = grow_dat$growth_t1,
               
               fert.N.obs = nrow(fert_dat),
               fert.plot = fert_dat$plot_int,
               fert.year = fert_dat$year_int,
               fert.PC1 = fert_dat$PC1,
               fert.PC2 = fert_dat$PC2,
               fert.PC3 = fert_dat$PC3,
               fert.size = fert_dat$standvol_t,
               fert.y = fert_dat$Goodbuds_t)
```

```{r mcmc setup}
inits<-function(){list(flow.mu=rnorm(1),
                       fert.mu=rnorm(1),
                       surv.mu=rnorm(1),
                       grow.mu=rnorm(1),
                       
                       flow.sigma.plot=rlnorm(1),
                       fert.sigma.plot=rlnorm(1),
                       surv.sigma.plot=rlnorm(1),
                       grow.sigma.plot=rlnorm(1),
                       
                       flow.sigma.year=rlnorm(1),
                       fert.sigma.year=rlnorm(1),
                       surv.sigma.year=rlnorm(1),
                       grow.sigma.year=rlnorm(1),
                       
                       fert.sigma.overdisp=rlnorm(1),
                       grow.sigma.eps=rlnorm(1),
                       
                       flow.z=matrix(rbinom(12,1,0.5),nrow=4,ncol=3),
                       fert.z=matrix(rbinom(12,1,0.5),nrow=4,ncol=3),
                       grow.z=matrix(rbinom(12,1,0.5),nrow=4,ncol=3),
                       surv.z=matrix(rbinom(12,1,0.5),nrow=4,ncol=3),
                       
                       flow.zsize=rbinom(1,1,0.5),
                       fert.zsize=rbinom(1,1,0.5),
                       grow.zsize=rbinom(1,1,0.5),
                       surv.zsize=rbinom(1,1,0.5)
                       )}

## Params to estimate
parameters<-c("flow.mu","fert.mu","surv.mu","grow.mu",
              "flow.bsize","fert.bsize","surv.bsize","grow.bsize",
              "flow.bclim","fert.bclim","surv.bclim","grow.bclim",
              "flow.z","fert.z","surv.z","grow.z",
              "flow.zsize","fert.zsize","surv.zsize","grow.zsize",
              "flow.sigma.plot","fert.sigma.plot","surv.sigma.plot","grow.sigma.plot",
              "flow.sigma.year","fert.sigma.year","surv.sigma.year","grow.sigma.year",
              "fert.sigma.overdisp",
              "grow.sigma.eps",
              "flow.fit","fert.fit","surv.fit","grow.fit",
              "flow.fit.new","fert.fit.new","surv.fit.new","grow.fit.new")

## MCMC settings
ni<-40000
nb<-10000
nt<-10
nc<-3

```

Run this mother-fucker.
```{r mcmc run}
allrates.out<-jags(data=cholla.dat,inits=inits,parameters.to.save=parameters,model.file="Bayes models\\HB_cholla_allrates_SVS.txt",
          n.thin=nt,n.chains=nc,n.burnin=nb,n.iter=ni,DIC=T,working.directory=getwd())
```

```{r diagnostics}
#mcmcplot(allrates.out)

par(mfrow=c(2,2))
plot(allrates.out$BUGSoutput$sims.list$flow.fit,
     allrates.out$BUGSoutput$sims.list$flow.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
plot(allrates.out$BUGSoutput$sims.list$fert.fit,
     allrates.out$BUGSoutput$sims.list$fert.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
plot(allrates.out$BUGSoutput$sims.list$grow.fit,
     allrates.out$BUGSoutput$sims.list$grow.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
plot(allrates.out$BUGSoutput$sims.list$surv.fit,
     allrates.out$BUGSoutput$sims.list$surv.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
```

```{r z scores}
allrates.out$BUGSoutput$mean$flow.z
allrates.out$BUGSoutput$mean$fert.z
allrates.out$BUGSoutput$mean$grow.z
allrates.out$BUGSoutput$mean$surv.z

## write z scores to file
write.csv(data.frame(allrates.out$BUGSoutput$summary),
          "allrates_SVS_out.csv")

```

Create a new "final" model based on these results for which coefficients are non-zero.
```{r model selection}
which(allrates.out$BUGSoutput$mean$flow.z>0.1,arr.ind=T)
which(allrates.out$BUGSoutput$mean$fert.z>0.1,arr.ind=T)
which(allrates.out$BUGSoutput$mean$grow.z>0.1,arr.ind=T)
which(allrates.out$BUGSoutput$mean$surv.z>0.1,arr.ind=T)

```

What about keeping size? Yes, keep all of them.
```{r}
allrates.out$BUGSoutput$mean$flow.zsize
allrates.out$BUGSoutput$mean$fert.zsize
allrates.out$BUGSoutput$mean$grow.zsize
allrates.out$BUGSoutput$mean$surv.zsize

```

Run the "selected model" only with coefficients that are likely (p > 0.1) non-zero.
```{r final model}
inits<-function(){list(flow.mu=rnorm(1,-10,2),
                       fert.mu=rnorm(1),
                       surv.mu=rnorm(1),
                       grow.mu=rnorm(1),
                       
                       flow.sigma.plot=rlnorm(1),
                       fert.sigma.plot=rlnorm(1),
                       surv.sigma.plot=rlnorm(1),
                       grow.sigma.plot=rlnorm(1),
                       
                       flow.sigma.year=rlnorm(1),
                       fert.sigma.year=rlnorm(1),
                       surv.sigma.year=rlnorm(1),
                       grow.sigma.year=rlnorm(1),
                       
                       fert.sigma.overdisp=rlnorm(1),
                       grow.sigma.eps=rlnorm(1),
                       
                       flow.bclim=matrix(rnorm(12),nrow=4,ncol=3),
                       fert.bclim=matrix(rnorm(12),nrow=4,ncol=3),
                       grow.bclim=matrix(rnorm(12),nrow=4,ncol=3),
                       surv.bclim=matrix(rnorm(12),nrow=4,ncol=3),
                       
                       flow.bsize=rnorm(1),
                       fert.bsize=rnorm(1),
                       grow.bsize=rnorm(1),
                       surv.bsize=rnorm(1)
                       )}

## Params to estimate
parameters<-c("flow.mu","fert.mu","surv.mu","grow.mu",
              "flow.bsize","fert.bsize","surv.bsize","grow.bsize",
              "flow.bclim","fert.bclim","surv.bclim","grow.bclim",
              "flow.sigma.plot","fert.sigma.plot","surv.sigma.plot","grow.sigma.plot",
              "flow.sigma.year","fert.sigma.year","surv.sigma.year","grow.sigma.year",
              "fert.sigma.overdisp",
              "grow.sigma.eps",
              "flow.fit","fert.fit","surv.fit","grow.fit",
              "flow.fit.new","fert.fit.new","surv.fit.new","grow.fit.new")

## MCMC settings
ni<-30000
nb<-5000
nt<-10
nc<-3

allrates.selected.out<-jags(data=cholla.dat,inits=inits,parameters.to.save=parameters,model.file="Bayes models\\HB_cholla_allrates_selected.txt",
          n.thin=nt,n.chains=nc,n.burnin=nb,n.iter=ni,DIC=T,working.directory=getwd())
```

```{r diagnostics}
#mcmcplot(allrates.selected.out)
par(mfrow=c(2,2))
plot(allrates.selected.out$BUGSoutput$sims.list$flow.fit,
     allrates.selected.out$BUGSoutput$sims.list$flow.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
plot(allrates.selected.out$BUGSoutput$sims.list$fert.fit,
     allrates.selected.out$BUGSoutput$sims.list$fert.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
plot(allrates.selected.out$BUGSoutput$sims.list$grow.fit,
     allrates.selected.out$BUGSoutput$sims.list$grow.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
plot(allrates.selected.out$BUGSoutput$sims.list$surv.fit,
     allrates.selected.out$BUGSoutput$sims.list$surv.fit.new,
     xlab="SSQ for actual data",
     ylab="SSQ for perfect (new) data")
abline(0,1, col='darkgray',lwd=3)
```

This all looks good. Write the mean parameter estimates, the summary, and a bunch of the posterior draws to file. 
```{r write final params}
write.csv(allrates.selected.out$BUGSoutput$summary,
          "allrates.selected.summary.csv")
saveRDS(allrates.selected.out$BUGSoutput$mean,
        "allrates.selected.mean.rds")

samples<-sample(1:7500,size=1000,replace=F)
allrates.selected.posterior<-data.frame(allrates.selected.out$BUGSoutput$sims.matrix[samples,])
write.csv(allrates.selected.posterior,
          "allrates.selected.posterior.csv")
```

Now I will pick up building graphics from these parameter estimates. If I want to run or revise these figures, I can just load the csv files rather than rerunning the full bayesian model (which takes several hours).

## Graphics
```{r read params}
## read in mean parameter estimates
mean_params <- readRDS("allrates.selected.mean.rds")
```

Start with defining the binning parameters for size. 
```{r binning params}
# there are the binning parameters for visualization
PC_bins <- 4
size_bin_num <- 3
bin_cols <- c("#c6dbef","#6baed6","#08519c")
alpha.col<-0.5

## here is how the size distn will be cut up:
size_cuts <- cholla.clim %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num)) %>% 
  group_by(size_bin) %>% 
  summarise(bin_mean = mean(standvol_t))
hist(cholla.clim$standvol_t)
abline(v = size_cuts$bin_mean)
```

Graphics for flowering:
```{r flowering graphics}

flow_dat_cuts <- flow_dat %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num),
         size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,Year_t) %>% 
  summarise(bin_mean = mean(standvol_t),
            mean_PC1 = mean(PC1),
            mean_PC2 = mean(PC2),
            mean_PC3 = mean(PC3),
            mean_flow = mean(Goodbuds_t > 0),
            bin_n = n())

#win.graph()
par(mfrow=c(1,3))
with(flow_dat_cuts,{
  plot(mean_PC1,mean_flow,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),lwd=1,
                         xlim=c(min(x_PC1),max(x_PC1)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC1,
        invlogit(mean_params$flow.mu + mean_params$flow.bsize*size_cuts$bin_mean[i] + 
                   mean_params$flow.bclim[1,1] * x_PC1),
        lwd=3,col=bin_cols[i])
  }

  plot(mean_PC2,mean_flow,col=alpha(bin_cols[size_bins],alpha.col),pch=16,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC2),max(x_PC2)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC2,
        invlogit(mean_params$flow.mu + mean_params$flow.bsize*size_cuts$bin_mean[i] + 
                   mean_params$flow.bclim[1,2] * x_PC2 + 
                   mean_params$flow.bclim[3,2] * x_PC2 * size_cuts$bin_mean[i] +
                   mean_params$flow.bclim[4,2] * (x_PC2^2) * size_cuts$bin_mean[i]),
        lwd=2,col=bin_cols[i])
  }
  
  plot(mean_PC3,mean_flow,bg=bin_cols[size_bins],pch=21,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC3),max(x_PC3)),ylim=c(0,1))
    for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC3,
        invlogit(mean_params$flow.mu + mean_params$flow.bsize*size_cuts$bin_mean[i] + 
                   mean_params$flow.bclim[1,3] * x_PC3 + 
                   mean_params$flow.bclim[3,3] * x_PC3 * size_cuts$bin_mean[i] +
                   mean_params$flow.bclim[4,3] * (x_PC3^2) * size_cuts$bin_mean[i]),
        lwd=3,col=bin_cols[i])
  }
})

```

Graphics for survival:
```{r survival graphics}

surv_dat_cuts <- surv_dat %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num),
         size_bins = as.integer(size_bin)) %>% 
  group_by(size_bins,Year_t) %>% 
  summarise(bin_mean = mean(standvol_t),
            mean_PC1 = mean(PC1),
            mean_PC2 = mean(PC2),
            mean_PC3 = mean(PC3),
            mean_surv = mean(Survival_t1),
            bin_n = n())

#win.graph()
par(mfrow=c(1,3))
with(surv_dat_cuts,{
  plot(mean_PC1,mean_surv,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC1),max(x_PC1)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC1,
        invlogit(mean_params$surv.mu + mean_params$surv.bsize*size_cuts$bin_mean[i] + 
                   mean_params$surv.bclim[1,1] * x_PC1),
        lwd=3,col=bin_cols[i])
  }

  plot(mean_PC2,mean_surv,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC2),max(x_PC2)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC2,
        invlogit(mean_params$surv.mu + mean_params$surv.bsize*size_cuts$bin_mean[i] + 
                   mean_params$surv.bclim[1,2] * x_PC2),
        lwd=3,col=bin_cols[i])
  }
  
  plot(mean_PC3,mean_surv,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC3),max(x_PC3)),ylim=c(0,1))
    for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC3,
        invlogit(mean_params$surv.mu + mean_params$surv.bsize*size_cuts$bin_mean[i] + 
                   mean_params$surv.bclim[1,3] * x_PC3 + 
                   mean_params$surv.bclim[2,3] * (x_PC3^2)),
        lwd=3,col=bin_cols[i])
  }
})

```

Graphics for growth:
```{r growth graphics}
grow_dat_cuts <- grow_dat %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num),
         size_bins = as.integer(size_bin)) 

#win.graph()
par(mfrow=c(1,3))
with(grow_dat_cuts,{
  plot(PC1,growth_t1,col=bin_cols[size_bins],pch=1,cex=1.5,
                         xlim=c(min(x_PC1),max(x_PC1)))
  plot(PC2,growth_t1,col=bin_cols[size_bins],pch=1,cex=1.5,
                         xlim=c(min(x_PC2),max(x_PC2)))
  plot(PC3,growth_t1,col=bin_cols[size_bins],pch=1,cex=1.5,
                         xlim=c(min(x_PC3),max(x_PC3)))

})

```


Graphics for fertility. This one is a little different because we have to chop the size distribution in a different way, since non-flowering plants are excluded from this data set. 
```{r fertility graphics}
fert_size_cuts <- cholla.clim %>% 
  filter(Goodbuds_t>0) %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num)) %>% 
  group_by(size_bin) %>% 
  summarise(bin_mean = mean(standvol_t))

fert_dat_cuts <- fert_dat %>% 
  mutate(size_bin = cut_number(standvol_t,size_bin_num),
         size_bins = as.integer(size_bin))

ymax.fert<-100

#win.graph()
par(mfrow=c(1,3))
with(fert_dat_cuts,{
  plot(PC1,Goodbuds_t,col=alpha(bin_cols[size_bins],alpha.fert),pch=1.4,cex=1.5,
                         xlim=c(min(x_PC1),max(x_PC1)),ylim=c(0,ymax.fert))

  plot(PC2,Goodbuds_t,col=alpha(bin_cols[size_bins],alpha.fert),pch=1.4,cex=1.5,
                         xlim=c(min(x_PC2),max(x_PC2)),ylim=c(0,ymax.fert))
  for(i in 1:length(fert_size_cuts$bin_mean)){
  lines(x_PC2,
        exp(mean_params$fert.mu + mean_params$fert.bsize*fert_size_cuts$bin_mean[i] + 
                   mean_params$fert.bclim[3,2] * x_PC2 * fert_size_cuts$bin_mean[i]),
        lwd=2,col=bin_cols[i])
  }
  
  plot(PC3,Goodbuds_t,col=alpha(bin_cols[size_bins],alpha.fert),pch=1.4,cex=1.5,
                         xlim=c(min(x_PC3),max(x_PC3)),ylim=c(0,ymax.fert))

})

```

## Big combo figure
Something like this will be in the final paper.

```{r combo fig}
par(mfrow=c(4,3))
with(flow_dat_cuts,{
  plot(mean_PC1,mean_flow,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),lwd=1,
                         xlim=c(min(x_PC1),max(x_PC1)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC1,
        invlogit(mean_params$flow.mu + mean_params$flow.bsize*size_cuts$bin_mean[i] + 
                   mean_params$flow.bclim[1,1] * x_PC1),
        lwd=3,col=bin_cols[i])
  }

  plot(mean_PC2,mean_flow,col=alpha(bin_cols[size_bins],alpha.col),pch=16,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC2),max(x_PC2)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC2,
        invlogit(mean_params$flow.mu + mean_params$flow.bsize*size_cuts$bin_mean[i] + 
                   mean_params$flow.bclim[1,2] * x_PC2 + 
                   mean_params$flow.bclim[3,2] * x_PC2 * size_cuts$bin_mean[i] +
                   mean_params$flow.bclim[4,2] * (x_PC2^2) * size_cuts$bin_mean[i]),
        lwd=2,col=bin_cols[i])
  }
  
  plot(mean_PC3,mean_flow,bg=bin_cols[size_bins],pch=21,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC3),max(x_PC3)),ylim=c(0,1))
    for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC3,
        invlogit(mean_params$flow.mu + mean_params$flow.bsize*size_cuts$bin_mean[i] + 
                   mean_params$flow.bclim[1,3] * x_PC3 + 
                   mean_params$flow.bclim[3,3] * x_PC3 * size_cuts$bin_mean[i] +
                   mean_params$flow.bclim[4,3] * (x_PC3^2) * size_cuts$bin_mean[i]),
        lwd=3,col=bin_cols[i])
  }
})

with(fert_dat_cuts,{
  plot(PC1,Goodbuds_t,col=alpha(bin_cols[size_bins],alpha.fert),pch=1.4,cex=1.5,
                         xlim=c(min(x_PC1),max(x_PC1)),ylim=c(0,ymax.fert))

  plot(PC2,Goodbuds_t,col=alpha(bin_cols[size_bins],alpha.fert),pch=1.4,cex=1.5,
                         xlim=c(min(x_PC2),max(x_PC2)),ylim=c(0,ymax.fert))
  for(i in 1:length(fert_size_cuts$bin_mean)){
  lines(x_PC2,
        exp(mean_params$fert.mu + mean_params$fert.bsize*fert_size_cuts$bin_mean[i] + 
                   mean_params$fert.bclim[3,2] * x_PC2 * fert_size_cuts$bin_mean[i]),
        lwd=2,col=bin_cols[i])
  }
  
  plot(PC3,Goodbuds_t,col=alpha(bin_cols[size_bins],alpha.fert),pch=1.4,cex=1.5,
                         xlim=c(min(x_PC3),max(x_PC3)),ylim=c(0,ymax.fert))

})

with(surv_dat_cuts,{
  plot(mean_PC1,mean_surv,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC1),max(x_PC1)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC1,
        invlogit(mean_params$surv.mu + mean_params$surv.bsize*size_cuts$bin_mean[i] + 
                   mean_params$surv.bclim[1,1] * x_PC1),
        lwd=3,col=bin_cols[i])
  }

  plot(mean_PC2,mean_surv,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC2),max(x_PC2)),ylim=c(0,1))
  for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC2,
        invlogit(mean_params$surv.mu + mean_params$surv.bsize*size_cuts$bin_mean[i] + 
                   mean_params$surv.bclim[1,2] * x_PC2),
        lwd=3,col=bin_cols[i])
  }
  
  plot(mean_PC3,mean_surv,col=bin_cols[size_bins],pch=1,cex= 3*(bin_n / max(bin_n)),
                         xlim=c(min(x_PC3),max(x_PC3)),ylim=c(0,1))
    for(i in 1:length(size_cuts$bin_mean)){
  lines(x_PC3,
        invlogit(mean_params$surv.mu + mean_params$surv.bsize*size_cuts$bin_mean[i] + 
                   mean_params$surv.bclim[1,3] * x_PC3 + 
                   mean_params$surv.bclim[2,3] * (x_PC3^2)),
        lwd=3,col=bin_cols[i])
  }
})

with(grow_dat_cuts,{
  plot(PC1,growth_t1,col=bin_cols[size_bins],pch=1,cex=1.5,
                         xlim=c(min(x_PC1),max(x_PC1)))
  plot(PC2,growth_t1,col=bin_cols[size_bins],pch=1,cex=1.5,
                         xlim=c(min(x_PC2),max(x_PC2)))
  plot(PC3,growth_t1,col=bin_cols[size_bins],pch=1,cex=1.5,
                         xlim=c(min(x_PC3),max(x_PC3)))

})
```

